---
title: "Customer Lifetime Value (CLV)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

I have been recently a lot about customer lifetime value and trying to understand the business decisions and math that drive it. To that end, I have found a lot of tutorials that try to explain it, but in order to learn for myself, I need to code and do the calculations. 


# Understanding clv

### Purpose: RFM and CLV analysis of customer:  

Goal 1: Segment the customer based on their purchase activity;  

Goal 2: Identify which type of customer should be promoted to make profit;  

Goal 3: Calculate the customer life time value  

### Step 1: Load data and check the data structure  

```{r dataLoad, message=FALSE}
library(tidyverse)
library(gridExtra)

## read in data
data_clv <- read_delim("C:/Users/akbar/Documents/Projects/Understanding_CLV/data/src/CDNOW_master.txt", delim = " ")

## data prep
data_prep <- data_clv %>%
     rename(id = " customer_id",
            date = " date",
            num = number_of_cds,
            amount = " dollar_value") %>%
     mutate(date = as.Date(as.character(date), "%Y%m%d"),
            id = as.numeric(id),
            num = as.numeric(num),
            amount = as.numeric(amount)) 

## number of unique customers
unique_cust <- data_prep %>%
     filter(!duplicated(id)) %>%
     .$id

glimpse(data_prep)

glimpse(unique_cust)
```

### Step 2: R,F,and M creation.
  
R(Recency): How recently did the customer purchase?  

F(Frequency): How often do they purchase?  

M(Monetary value): How much do they spend each time on average?

```{r RFM_analysis}
startDate <- min(data_prep$date)
endDate <- as.Date("1998-01-01")

## cutting data down to date ranges set above
temp_1 <- data_prep %>%
     arrange(desc(date)) %>%
     filter(date >= startDate & date <= endDate)

## making id a factor so that we can count the number of occurences
temp <- temp_1 %>%
     mutate(id = as.factor(id))

temp_f <- temp_1 %>%
  group_by(id) %>%
  summarise(frequency = n())

## creating a temporary place holder for M
temp_m <- temp_1 %>%
     group_by(id) %>%
     summarise(monetary = sum(amount))

## creating R, F and final M, based on F
temp_2 <- temp_1 %>%
     filter(!duplicated(id)) %>%
     left_join(temp_f) %>%
     left_join(temp_m) %>%
     mutate(recency = as.numeric(difftime(endDate, date, units = "days")),
            monetary = round(monetary/frequency, 2)) %>%
     filter(monetary < 600) %>%
     arrange(id) %>%
     select(id, recency, frequency, monetary)

glimpse(temp_2)
```

### Step 3: RFM rate

```{r rfm_rate}
## Creating the rates for R, F, M and combined RFM value
temp_fin <- temp_2 %>%
     mutate(rec_rate = as.numeric(cut_interval(recency, 5)),
            frq_rate = as.numeric(cut_interval(frequency, 5)),
            mon_rate = as.numeric(cut_interval(monetary, 5))) %>%
     select(id, recency, frequency, monetary, rec_rate, frq_rate, mon_rate) %>%
     mutate(rfm_score = as.numeric(paste0(rec_rate, frq_rate, mon_rate)))

```

### https://www.r-bloggers.com/rfm-customer-analysis-with-r-language/

```{r}

```

